Classical Planning Algorithms ¡ on the Atari Video Games

### 2014-11-19 icaps15-paper/paper.tex
added new games

\newcommand{\mc}[2]{\multicolumn{#1}{b}{#2}}

@@

\newcommand{\numgames}{$35$}

\newcommand{\Assault}{\propername{Assault}}
\newcommand{\Asterix}{\propername{Asterix}}
\newcommand{\Asteroids}{\propername{Asteroids}}
\newcommand{\Atlantis}{\propername{Atlantis}}
\newcommand{\BankHeist}{\propername{Bank Heist}}
\newcommand{\BattleZone}{\propername{Battle Zone}}
\newcommand{\BeamRider}{\propername{Beam Rider}}
\newcommand{\Berzerk}{\propername{Berzerk}}
\newcommand{\Bowling}{\propername{Bowling}}
\newcommand{\Boxing}{\propername{Boxing}}
\newcommand{\Breakout}{\propername{Breakout}}
\newcommand{\Centipede}{\propername{Centipede}}
\newcommand{\ChopperCommand}{\propername{Chopper Command}}
\newcommand{\CrazyClimber}{\propername{Crazy Climber}}
\newcommand{\Enduro}{\propername{Enduro}}
\newcommand{\Freeway}{\propername{Freeway}}
\newcommand{\Frostbite}{\propername{Frostbite}}
\newcommand{\Gravitar}{\propername{Gravitar}}
\newcommand{\Hero}{\propername{Hero}}
\newcommand{\JamesBond}{\propername{James Bond}}
\newcommand{\Kangaroo}{\propername{Kangaroo}}
\newcommand{\Krull}{\propername{Krull}}
\newcommand{\Montezuma}{\propername{Montezuma Revenge}}
\newcommand{\MsPacman}{\propername{Ms Pacman}}
\newcommand{\Pong}{\propername{Pong}}
\newcommand{\Pooyan}{\propername{Pooyan}}
\newcommand{\Qbert}{\propername{Q*Bert}}
\newcommand{\RiverRaid}{\propername{Riverraid}}
\newcommand{\Seaquest}{\propername{Seaquest}}
\newcommand{\SpaceInvaders}{\propername{Space Invaders}}
\newcommand{\StarGunner}{\propername{Star Gunner}}
\newcommand{\Tennis}{\propername{Tennis}}
\newcommand{\Venture}{\propername{Venture}}
\newcommand{\VideoPinball}{\propername{Video Pinball}}
\newcommand{\Zaxxon}{\propername{Zaxxon}}

@@

Last, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. We hope to investigate the implications of
this gap in the future.

@@

More interestingly,  \IW(1) outscores \UCT\  in $17$ of the \numgames\ games,
while \BFS\ outscores \UCT\ in $18$.
%\footnote{These numbers can actually
%improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $9$ and $11$ games respectively.
The relative performance between $\IW(1)$ and \BFS\ is closer with $\IW(1)$ being  the
best of the two in $12$ games, and  \BFS\  in $11$. In terms of the number of games 
where an algorithm is the overall winner, the situation is  even: 
$\IW(1)$ is the overall best in $12$ games, \BFS\ in $11$ games, and \UCT\ in $12$ games. 
Also, \BRFS\ is best in $1$ game (\Centipede), while the other three algorithms
are tied in another game (\Pong).

@@

Likewise, in \Freeway\ and \Berzerk\ both $\IW(1)$ and \BFS\ attain a better
score than the baseline semi-random algorithm \emph{Perturb} in
\cite{bellemare:jair2013}, that beats \UCT\ on those games.
%% to say the 2 lines below must add context: .. eg
%%% ``which in *these games* outperforms both \UCT\ and the reinforcement learning methods.''
%% add it if that's the case ..
\emph{Perturb} is a simple algorithm
that selects  a fixed  action with probability $0.95$, and a random
action with probability $0.05$.  For \emph{Perturb}, BNVB do not report 
the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
are deep but can be reached by  repeating the same action.
This is the case of \Freeway, where a chicken has to run across a ten lane
highway filled with traffic. Every time the chicken gets across, there is one
unit of reward. If the chicken is hit by a car, it has to start again. 
In \Freeway, only $12$ out of $18$ possible actions have an effect:
$6$ actions in moving the chicken up one lane (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
$6$ in moving the chicken down one lane (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
when the selected fixed action is one of the $6$ that moves the chicken up.  As noted 
in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
across the highway at all. The reason for \UCT\ to perform erratically 
is that the roll-outs do not manage to collect any reward. A sequence of actions 
with at least $48$~\footnote{One needs to move up the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
highway.} actions embedded that move the chicken up are required in order 
to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
that \emph{also} avoids incoming traffic in a timely fashion is rather low.
%% the previous number was not correct; no-op actions do not count, and thus needs
%%% 200 more move up actions than move down actions, although this is not accurate either
%%%% as move down actions that take chicken below the screen count as no-ops too ..
$\IW(1)$ does not have this limitation and is best in \Freeway.

>>>>

\newcommand{\mc}[2]{\multicolumn{#1}{a}{#2}}

@@

\newcommand{\numgames}{$48$}

\newcommand{\Assault}{\propername{Assault}}
\newcommand{\Asterix}{\propername{Asterix}}
\newcommand{\Asteroids}{\propername{Asteroids}}
\newcommand{\Atlantis}{\propername{Atlantis}}
\newcommand{\BankHeist}{\propername{Bank Heist}}
\newcommand{\BattleZone}{\propername{Battle Zone}}
\newcommand{\BeamRider}{\propername{Beam Rider}}
\newcommand{\Berzerk}{\propername{Berzerk}}
\newcommand{\Bowling}{\propername{Bowling}}
\newcommand{\Boxing}{\propername{Boxing}}
\newcommand{\Breakout}{\propername{Breakout}}
\newcommand{\Carnival}{\propername{Carnival}}
\newcommand{\Centipede}{\propername{Centipede}}
\newcommand{\ChopperCommand}{\propername{Chopper Command}}
\newcommand{\CrazyClimber}{\propername{Crazy Climber}}
\newcommand{\DoubleDunk}{\propername{Double Dunk}}
\newcommand{\Enduro}{\propername{Enduro}}
\newcommand{\FishingDerby}{\propername{Fishing Derby}}
\newcommand{\Freeway}{\propername{Freeway}}
\newcommand{\Frostbite}{\propername{Frostbite}}
\newcommand{\Gravitar}{\propername{Gravitar}}
\newcommand{\Hero}{\propername{Hero}}
\newcommand{\JamesBond}{\propername{James Bond}}
\newcommand{\JourneyEscape}{\propername{Journey Escape}}
\newcommand{\Kangaroo}{\propername{Kangaroo}}
\newcommand{\Krull}{\propername{Krull}}
\newcommand{\KungFuMaster}{\propername{Kung Fu Master}}
\newcommand{\Montezuma}{\propername{Montezuma Revenge}}
\newcommand{\MsPacman}{\propername{Ms Pacman}}
\newcommand{\NameThisGame}{\propername{Name This Game}}
\newcommand{\Pong}{\propername{Pong}}
\newcommand{\Pooyan}{\propername{Pooyan}}
\newcommand{\PrivateEye}{\propername{Private Eye}}
\newcommand{\Qbert}{\propername{Q*Bert}}
\newcommand{\RiverRaid}{\propername{Riverraid}}
\newcommand{\RobotTank}{\propername{Robot Tank}}
\newcommand{\Seaquest}{\propername{Seaquest}}
\newcommand{\SpaceInvaders}{\propername{Space Invaders}}
\newcommand{\StarGunner}{\propername{Star Gunner}}
\newcommand{\Tennis}{\propername{Tennis}}
\newcommand{\TimePilot}{\propername{Time Pilot}}
\newcommand{\Tutankham}{\propername{Tutankham}}
\newcommand{\UpNDown}{\propername{Up And Down}}
\newcommand{\Venture}{\propername{Venture}}
\newcommand{\VideoPinball}{\propername{Video Pinball}}
\newcommand{\WizardOfWor}{\propername{Wizard Of Wor}}
\newcommand{\Zaxxon}{\propername{Zaxxon}}

@@

Last, \IW(1) and \BFS\  in the lookahead select an action every $5$ frames, while \UCT\  selects 
an action in every frame.  This means that in order to explore a branch 
$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
get to choose $60$ actions, both however using the same $300$ frames from the
budget. For this, we  followed the setup  of \BRFS\ in BNVB 
that also selects actions every $5$ frames, matching the behavior
of the emulator that requests an action also every $5$ frames. Since the 
lookahead budget is given by a maximum number of (new) frames, and the time 
is mostly taken by calls to the emulator, this may not be the best choice  
for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
options afforded by the budget. We hope to investigate the implications of
this gap in the future.

@@

More interestingly,  \IW(1) outscores \UCT\  in $27$ of the \numgames\ games,
while \BFS\ outscores \UCT\ in $21$.
%\footnote{These numbers can actually
%improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $15$ and $17$ games respectively.
The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
best of the two in $22$ games, and  \BFS\  in $12$. In terms of the number of games 
where an algorithm is the overall winner, $\IW(1)$  domainates the others: 
$\IW(1)$ is the overall best in $23$ games, \BFS\ in $11$ games, and \UCT\ in $18$ games. 
Also, \BRFS\ is best in $2$ games (\Centipede, tied up in \Boxing), while the other three algorithms
are tied in another 2 games (\Pong, \Boxing), and $\IW(1)$ and \BFS\ are tied in \TimePilot.

@@

Likewise, in \Freeway\ and \Berzerk\ both $\IW(1)$ and \BFS\ attain a better
score than the baseline semi-random algorithm \emph{Perturb} in
\cite{bellemare:jair2013}, that beats \UCT\ on those games.
%% to say the 2 lines below must add context: .. eg
%%% ``which in *these games* outperforms both \UCT\ and the reinforcement learning methods.''
%% add it if that's the case ..
\emph{Perturb} is a simple algorithm
that selects  a fixed  action with probability $0.95$, and a random
action with probability $0.05$.  For \emph{Perturb}, BNVB do not report 
the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
are deep but can be reached by  repeating the same action.
This is the case of \Freeway, where a chicken has to run across a ten lane
highway filled with traffic. Every time the chicken gets across, there is one
unit of reward. If the chicken is hit by a car, it goes back some lanes. 
In \Freeway, only $12$ out of $18$ possible actions have an effect:
$6$ actions in moving the chicken up one lane (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
$6$ in moving the chicken down one lane (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
when the selected fixed action is one of the $6$ that moves the chicken up.  As noted 
in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
across the highway at all. The reason for \UCT\ to perform erratically 
is that the roll-outs do not manage to collect any reward. A sequence of actions 
with at least $48$~\footnote{One needs to move up the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
highway.} actions embedded that move the chicken up are required in order 
to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
that \emph{also} avoids incoming traffic in a timely fashion is rather low.
%% the previous number was not correct; no-op actions do not count, and thus needs
%%% 200 more move up actions than move down actions, although this is not accurate either
%%%% as move down actions that take chicken below the screen count as no-ops too ..
$\IW(1)$ does not have this limitation and is best in \Freeway.
